\chapter{Summary and Conclusion}

\acrfull{abe} enforces flexible access control and thus offers great potential for many use cases.
In this thesis, I presented \acrshort{abe} and evaluated its practical feasibility on constrained \acrshort{iot} nodes.
More specifically, I implemented an \acrshort{abe} library with two schemes and evaluated it on an ARM Cortex-M4 SoC.
To this end, I modified the underlying elliptic curve and pairing library \texttt{rabe-bn} to run without dynamic memory allocation.
For the evaluation of the library, all four \acrshort{abe} algorithms (Encrypt, Decrypt, Setup and KeyGen) were tested on the embedded SoC.
Runtimes, RAM use and binary sizes were recorded.

Computing the bilinear pairings underlying most \acrshortpl{abes} on the SoC is feasible, but comes at a rather high cost.
Comparison with other pairing implementations shows that there is some room for improvement here, but the the operations remain expensive.
In addition to computation time, memory size is a major bottleneck for pairing-based implementations.
% The GPSW scheme is limited to small and medium policies for key generation and decryption. 
% With less memory, even computation of a single pairing fails because of insufficient RAM.

In the use case presented in the introduction, the constrained node only needs to encrypt data with \acrshort{abe}.
This is feasible, but it does take up a significantly portion of the available resources.
The runtime is in the order of a few seconds and only increases linearly as more attributes are added.
For use cases where such a delay is not practical, re-use of the encapsulated symmetric key might mitigate the issue.

With decryption, the case is different: 
No problems were encountered with the pairing-free YCT scheme.
With the pairing-based GPSW scheme, however, encryption failed when using \glspl{access-tree} with more than two levels due to too little RAM.
Also, runtimes were much longer than for encryption and increased much faster with growing policy size.
While very long runtimes might sometimes be acceptable, the RAM limitation poses a hard limit.

If decryption with large policies is necessary on the SoC, it is therefore advisable to choose the pairing-free YCT scheme.
The security of \acrshort{abe} without pairings, however, remains questionable, see~\cite{herranz_attacking_2020}.
Further research in this area is needed.\\

In short, \acrlong{abe} is feasible on the considered hardware, especially if only encryption is needed and the number of attributes is not too large.
Still, it comes at a high cost in runtimes and memory consumption.
Especially decryption with the pairing-based scheme is not feasible because runtimes are very long and only \glspl{access-tree} with few levels can be used.